---
title: "Practical Machine Learning: Course Project"

output: html_document
---

###Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self-movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset).

###Data
The training data for this project are available here: <https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv><br />
The test data are available here: <https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv><br />
The data for this project come from this source: <http://groupware.les.inf.puc-rio.br/har>. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

###Project Goal
The goal of this project is to predict the manner in which the people did the exercises, which is defined in the “classe” variable in the training dataset. The goal is also describing how the prediction model is built, how it is cross validated, evaluation of the expected out of sample error, and explaining the reasons of the choices made to build this model. The prediction model will be used to predict 20 different test cases.

###Libraries

```{r results='hide', message=FALSE, warning=FALSE}
library(caret)
library(randomForest)
library(rpart)
```

###Data Loading

```{r}
#Load training_data and testing_data, and than replace invalid strings as NA
training_data <- read.csv("pml-training.csv", na.strings = c("NA", "#DIV/0!", ""))
testing_data  <- read.csv("pml-testing.csv",  na.strings = c("NA", "#DIV/0!", ""))

dim(training_data)
dim(testing_data)
```

###Processing Data

```{r}
#Delete columns with NA in testing_data
training_data <- training_data[, colSums(is.na(testing_data)) == 0]
testing_data  <- testing_data[, colSums(is.na(testing_data)) == 0]

#Delete some irrelevant variables: user_name, raw_timestamp_part_1, raw_timestamp_part_,2 cvtd_timestamp, new_window, num_window
training_data  <- training_data[, -c(1:7)]
testing_data   <- testing_data[, -c(1:7)]

dim(training_data)
dim(testing_data)
```

###Cross Validation

```{r}
#Divide training_data to sub_training_data and sub_testing_data (75% and 25% resp.)
part_ind           <- createDataPartition(y = training_data$classe, p = 0.75, list = FALSE,)
sub_training_data  <- training_data[part_ind,]
sub_testing_data   <- training_data[-part_ind,]
```

###Create Random Forest and Decision Tree models. Select the best.

```{r}
#Random Forest Model
rand_forest_model <- randomForest(classe ~. , data = sub_training_data, method = "class")
pred_rand_forest  <- predict(rand_forest_model, sub_testing_data, type = "class")
res_rand_forest    <- confusionMatrix(pred_rand_forest, sub_testing_data$classe)

#Decision Tree Model
dec_tree_model  <- rpart(classe ~ ., data = sub_training_data, method = "class")
pred_dec_tree   <- predict(dec_tree_model, sub_testing_data, type = "class")
res_rand_tree   <- confusionMatrix(pred_dec_tree, sub_testing_data$classe)

#Compare models and choose with the best accuracy (random forest)
res_comp <- data.frame(res_rand_forest$overall, res_rand_tree$overall)
res_comp
```

###Submission
```{r}
#Predict outcome levels on the testing_data using random forest model
final_pred <- predict(rand_forest_model, testing_data, type = "class")
final_pred
```